An agent-based model of local food systems

### 2016-05-13 draft.tex
update

\begin{abstract}
\noindent 
It seems that local food systems are driven by factors that play minor roles in conventional food systems, which are characterized in terms of interaction and diversity. Key interactions occur among consumers, exchanging information, and between consumers and producers, building trust. Diverse motives for food transaction include social, cultural, political, and ethical considerations. In order to understand local food systems, given the dynamical process and heterogeneity, agent-based computational modeling is considered more suitable than the traditional analytical approach.\\\\
\noindent
Keywords: \textit{local food system, ABM}
\end{abstract}

@@

\subsection{Agents}
There are two types of agents: consumer $(i \in \{1,...,I\})$ and producer $(j \in \{1,...,J\})$. At each period, a typical consumer chooses a single producer by comparing the utility obtained upon purchasing food from each producer. The $i$th consumer's utility function of the purchase from the $j$th producer is defined as follows:
\begin{equation} \label{util}
  u_{ij} = \alpha_i t_{ij} + \beta_i e_j - \gamma_i d_{ij} - p_j .
\end{equation}
Through adjusting the price, each producer ``tries to'' maximize the profit:
\begin{equation}  \label{profit}
  \pi_j = p_j Q_j - c_j - k e_j .
\end{equation}
We stress that the producer's maximization is only an attempt due to the bounded rationality. Vendors at a farmers' market, for example, only incorporate neighboring competitors' prices or a limited amount of marketing intelligence. In the next section, we explain the variables and parameters.

@@

\subsubsection{Trust update}
At each period, a consumer updates her trust levels in all the producers. Here, as well, we try to incorporate social embeddedness into the process. The idea is that the more embedded, the greater chance to interact with the producer and therefore increase the trust. As a result, we assume that an upward adjustment takes place with probability of $0.5 + e_j \slash 2$. Specifically, the $i$th consumer's trust in $j$th producer is updated as follows:
\begin{displaymath}
t'_{ij} =
\begin{cases}
t_{ij} \times 1.05 & \text{with probability } 0.5 + e_j \slash 2\\
t_{ij} \times 0.99 & \text{otherwise}
\end{cases}
\end{displaymath}
The mild downward adjustment reflects gradual ``forgetting''. Finally, there is no trust update for conventional producers.

@@

\subsubsection{Networking \& Recommendation}
Given the socially embedded nature, connecting consumers and producers is often one of the main objectives of local food systems \parencite{Martinez2010}. The modeling questions are: What is the network for? and How is it formed? It seems true that social interaction in itself is an end, and therefore we may add a connectivity measure to the utility function. Here we see the connections between consumers as a friendship network, over which information about producers is exchanged. As explained above, the consumer's trust in a producer increases only when she makes a purchase, yet trust level is a key determinant of the consumer's decision. Realistically, however, consumers learn about new producers and try them out every now and then. One source of such information is those who participate in the same farmers' market or CSA. In essence, food acts as a medium for interaction, and a seller facilitate network formation among the customers.
%(We regard a collection of individual producers as an institution called a seller.)
\begin{figure}[h]
\centering
  \includegraphics[scale=0.9]{img02}
  \caption{Seller initiated network}
\end{figure}

To implement the idea, first we use a $I \times I$ adjacency matrix \parencite{Jackson2008} to represent the network, which is assumed to be undirected and have zero for all the initial entry. Note that a seller is absent in the permanent network as its role is only to temporarily facilitate interaction among customers at each period. Second, we assume that upon each purchase a consumer has a random chance to communicate and form a tie with another customer in the same seller initiated network. It is random because people do not communicate strictly every single time. The probability is equal to the producer's embeddedness level due to the natural correspondence between embeddedness and chance for social interaction. As a result, if a consumer buys from a more embedded producer, she has a greater chance for connection, but it does not mean a new connection as it may be with an existing friend. Once a tie is formed, there is no decay over the subsequent periods. Finally, to specify the mechanism of query and recommendation about a new producer, we have the following story:
\begin{adjustwidth}{1cm}{}
\textit{This week, as usual, after examining all the sellers, she chooses the regular one (trusted and close yet a bit pricy). But, she wonders how a producer sitting at the third place in her ranking might be. She has never tried his food but the price looks good...}
\end{adjustwidth}
So, each consumer sends a query about a new producer, who gives the highest utility among all the new producers. Friends, if any, report back their own trust levels in that producer. She aggregates the information and adopts it as her own trust in the new producer. Specifically, she adopts 90\% of the median of the friends' trust levels.
%Create opportunities for new producers and interesting dynamics

\section{Discussion}
\subsection{Outputs}
Figure 3 shows a typical output of the model, where the graph traces the number of local producers over 156 periods. Though far from complete, we try to define the units of the variables and calibrate the parameters on the basis of 1 period = 1 week. Hence, 156 periods are meant to be 3 years.
\begin{figure}[h]
\centering
  \includegraphics[scale=0.4]{img03}
  \caption{A typical output}
\end{figure}

\noindent At this point, there is not much to say about particular outputs, but the underlying mechanics of the model could work. The philosophy of ABM is summarized by what \textcite{Epstein1996} call ``generative'' social science. Before drawing any conclusion, therefore, we must look at data, figure out empirical regularities, apply alternative mechanism of interactions, and calibrate parameters in order to generate as much the empirical phenomena as possible. In what follows, we present a long list of work ahead.

>>>>

\begin{abstract}
\noindent 
It seems that local food systems are driven by the factors that play minor roles in conventional food systems, which are characterized by interaction and individual heterogeneity. There are two key interactions: networking \& recommendation among consumers, and trust building between consumers and producers. Diverse motives for food transaction include social, cultural, political, and ethical considerations. In order to understand local food systems, given the dynamical process and heterogeneity, agent-based computational modeling is considered more suitable than the traditional analytical approach.\\\\
\noindent
Keywords: \textit{local food system, ABM}
\end{abstract}

@@

\subsection{Agents}
There are two types of agents: consumer $i \in \{1,...,I\}$ and producer $j \in \{1,...,J\}$. At each period, a typical consumer chooses a single producer by comparing the utility obtained upon purchasing food from each producer. The $i$th consumer's utility function of the purchase from the $j$th producer is defined as follows:
\begin{equation} \label{util}
  u_{ij} = \alpha_i t_{ij} + \beta_i e_j - \gamma_i d_{ij} - p_j .
\end{equation}
Through adjusting the price, each producer ``tries to'' maximize the profit:
\begin{equation}  \label{profit}
  \pi_j = p_j Q_j - c_j - k e_j .
\end{equation}
We stress that the producer's maximization is only an attempt due to the bounded rationality. Vendors at a farmers' market, for example, only incorporate neighboring competitors' prices or a limited amount of marketing intelligence. In the next section, we explain the variables and parameters.

@@

\subsubsection{Trust update}
At each period, a consumer updates her trust levels in all the producers. Here, as well, we try to incorporate social embeddedness into the process. The idea is that the more embedded, the greater chance to interact with the producer and therefore increase the trust. As a result, we assume that an upward adjustment takes place with probability of $0.5 + \nicefrac{e_j}{2}$. Specifically, the $i$th consumer's trust in $j$th producer is updated as follows:
\begin{displaymath}
t'_{ij} =
\begin{cases}
t_{ij} \times 1.05 & \text{with probability } 0.5 + \dfrac{e_j}{2}\\
t_{ij} \times 0.99 & \text{otherwise}
\end{cases}
\end{displaymath}
The mild downward adjustment reflects gradual ``forgetting''. Finally, there is no trust update for conventional producers.

@@

\subsubsection{Networking \& Recommendation}
Given the socially embedded nature, connecting consumers and producers is often one of the main objectives of local food systems \parencite{Martinez2010}. The modeling questions are: What is the network for? and How is it formed? It seems true that social interaction in itself is an end, and therefore we may add a connectivity measure to the utility function. Here we see the connections between consumers as a friendship network, over which information about producers is exchanged. As explained above, the consumer's trust in a producer increases only when she makes a purchase, yet trust level is a key determinant of the consumer's decision. Realistically, however, consumers learn about new producers and try them out every now and then. One source of such information is those who participate in the same farmers' market or CSA. In essence, food acts as a medium for interaction, and a seller facilitates network formation among the customers. (n.b. In this section, by `new' producers, we mean those in which a consumer has the lowest trust level.)
%(We regard a collection of individual producers as an institution called a seller.)
\begin{figure}[h]
\centering
  \includegraphics[scale=0.9]{img02}
  \caption{Temporary seller initiated network}
\end{figure}

To implement the idea, first we use an $I \times I$ adjacency matrix \parencite{Jackson2008} to represent the network, which is assumed to be undirected and have zero for all the initial entry. Note that a seller is absent in the permanent network as its role is only to temporarily facilitate interaction among customers at each period. Second, we assume that upon each purchase a consumer has a random chance to communicate and form a tie with another customer in the same seller initiated network. It is random because people do not communicate strictly every single time. The probability is equal to the producer's embeddedness level due to the natural correspondence between embeddedness level and chance for social interaction. Thus, if a consumer buys from a more embedded producer, she has a greater chance for connection, but it does not necessarily mean a new connection as it may be with an existing friend. Once a tie is formed, there is no decay over the subsequent periods. Finally, to specify the mechanism of query and recommendation about a new producer, we have the following story:
\begin{adjustwidth}{1cm}{}
\textit{This week, as usual, after examining all the sellers, she chooses the regular one (trusted and close yet a bit pricy). But, she wonders how a producer sitting at the third place in her ranking might be. She has never tried his food but the price looks good...}
\end{adjustwidth}
So, each consumer sends a query about a new producer, who gives the highest utility among all the new producers. Friends, if any, report back their own trust levels in that producer. She aggregates the information and adopts it as her own trust in the new producer. Specifically, she adopts 90\% of the median of the friends' trust levels.
%Create opportunities for new producers and interesting dynamics

\section{Discussion}
\subsection{Outputs}
Figure 3 shows a typical output of the model, where the graph traces the number of local producers over 156 periods. Though far from complete, we try to define the units of the variables and calibrate the parameters on the basis of 1 period = 1 week. Hence, 156 periods are meant to be 3 years.
\begin{figure}[h]
\centering
  \includegraphics[scale=0.4]{img03}
  \caption{A typical output}
\end{figure}
At this point, there is not much to say about particular outputs, but the underlying mechanics of the model could work. The philosophy of ABM is summarized by what \textcite{Epstein1996} call ``generative'' social science. Before drawing any conclusion, therefore, we must look at data, figure out empirical regularities, apply alternative mechanism of interactions, and calibrate parameters in order to generate as much the empirical phenomena as possible. In what follows, we present a long list of work ahead.
